### 2020-12-14 exploratory/2020-11_Nov-Dec/plots.Rmd
More plots with debugging split fraction mapped

```{r load_opticlust}
opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf()
sum_opticlust <- opticlust %>% 
  group_by(dataset) %>% 
  summarize(mean_mcc = mean(mcc),
            sd_mcc = sd(mcc),
            mean_sec = mean(sec),
            sd_sec = sd(sec))#,
            #mean_shannon = mean(shannon))
```
```{r mcc_opticlust}
opticlust %>% 
  ggplot(aes(dataset, mcc)) +
  geom_jitter(alpha = 0.5) +
  ylim(0.5, 1)
```
```{r runtime_opticlust}
opticlust %>% 
  ggplot(aes(num_seqs, sec, color = dataset)) +
  geom_point(alpha = 0.5)
```
## fit to reference databases

@@

```{r fraction-mapped_fit-db}
optifit_dbs %>% filter(method == 'closed') %>% 
  ggplot(aes(x=dataset, y=fraction_mapped, color=ref)) +
  geom_boxplot() + 
  facet_wrap('method') +
  ylim(0, 1) +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```

Plotted the fraction mapped for open-reference clustering too as a sanity check.
Final plots to be included in the paper won't include open-ref.

```{r runtime_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=sec, color=ref)) +
  geom_boxplot(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_sec), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```
```{r diversity_fit-db, eval=FALSE}
optifit_dbs %>% 
  ggplot(aes(x=method, y=shannon, color=ref)) +
  geom_jitter(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_shannon), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

@@

```{r fraction-mapped-grid_fit-split}
optifit_split %>% 
  filter(tool == 'mothur', is.na(ref)) %>% 
  filter(dataset == 'soil') %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_point(alpha = 0.5, size=1) +
  facet_grid(dataset ~ method) +
  ylim(0, 1) +
  labs(title="Sequences mapped during open-reference OptiFit",
       x='reference fraction',
       y='fraction mapped')
```

@@

```{r mcc_all-opti}
sum_opti_all %>% 
  ggplot(aes(strategy, mcc_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ .) +
  scale_color_manual(values = tri_colors) +
  ylim(0, 1) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_flip() +
  labs(title = 'OTU Quality', x = '', y = 'Median MCC')
```
```{r runtime_all-opti}
sum_opti_all %>% 
  ggplot(aes(strategy, sec_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
```{r memory_all-opti}
sum_opti_all %>% 
  ggplot(aes(strategy, mem_gb_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r metrics_all-human}
all_human %>% 
  pivot_longer(c(mcc, sec, mem_gb, fraction_mapped), names_to = 'metric') %>% 
  ggplot(aes(strategy, value, color = tool)) +
  geom_boxplot() +
  facet_wrap('metric', scales = 'free') +
  labs(x='', y='',  
       caption = 'Only the human dataset is shown here. The patterns are similar with other datasets.')
```

>>>>

```{r load_opticlust}
opticlust <- read_tsv(here('subworkflows/1_prep_samples/results/opticlust_results.tsv')) %>% 
  full_join(read_tsv(here('subworkflows/1_prep_samples/results/dataset_sizes.tsv'))) %>% 
  mutate_perf()
sum_opticlust <- opticlust %>% 
  group_by(dataset) %>% 
  summarize(mean_mcc = mean(mcc),
            sd_mcc = sd(mcc),
            mean_sec = mean(sec),
            sd_sec = sd(sec),
            mean_mem_gb = mean(mem_gb))#,
            #mean_shannon = mean(shannon))
```
```{r mcc_opticlust}
opticlust %>% 
  ggplot(aes(dataset, mcc)) +
  geom_jitter(alpha = 0.5) +
  ylim(0.5, 1)
```
```{r runtime_opticlust}
opticlust %>% 
  ggplot(aes(num_seqs, sec, color = dataset)) +
  geom_point(alpha = 0.5)
```
## fit to reference databases

@@

```{r fraction-mapped_fit-db}
optifit_dbs %>% filter(method == 'closed') %>% 
  ggplot(aes(x=dataset, y=fraction_mapped, color=ref)) +
  geom_boxplot() + 
  facet_wrap('method') +
  ylim(0, 1) +
  scale_color_manual(values=tri_colors) +
  labs(title="Sequences mapped during closed-reference OptiFit") +
  theme(plot.caption = element_markdown())
```

```{r runtime_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=sec, color=ref)) +
  geom_boxplot(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_sec), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r memory_fit-db}
optifit_dbs %>% 
  ggplot(aes(x=method, y=mem_gb, color=ref)) +
  geom_boxplot(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_mem_gb), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

```{r diversity_fit-db, eval=FALSE}
optifit_dbs %>% 
  ggplot(aes(x=method, y=shannon, color=ref)) +
  geom_jitter(alpha = 0.5) +
  geom_hline(aes(yintercept = mean_shannon), sum_opticlust) +
  facet_wrap('dataset', scales = 'free') +
  scale_color_manual(values=tri_colors) +
  labs(title='OptiFit runtime with reference databases',
       caption='Black line: _de novo_ clustering') +
  theme(plot.caption = element_markdown())
```

@@

```{r fraction-mapped-grid_fit-split}
optifit_split %>% 
  filter(tool == 'mothur', is.na(ref)) %>% 
  filter(dataset == 'soil') %>% 
  ggplot(aes(x=ref_frac, y=fraction_mapped, color=ref_weight)) +
  geom_point(alpha = 0.5, size=1) +
  facet_grid(dataset ~ method) +
  ylim(0, 1) +
  labs(title="Sequences mapped during OptiFit",
       x='reference fraction',
       y='fraction mapped',
       caption=glue("Ran with {optifit_split %>% pull(seed) %>% unique() %>% length()} seeds"))
```

@@

```{r mcc_all-opti, fig.width=10, fig.height=3.5}
sum_opti_all %>% 
  ggplot(aes(strategy, mcc_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., nrow=1) +
  scale_color_manual(values = tri_colors) +
  ylim(0, 1) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  coord_flip() +
  labs(title = 'OTU Quality', x = '', y = 'Median MCC')
```
```{r runtime_all-opti}
sum_opti_all %>% 
  ggplot(aes(strategy, sec_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
```{r memory_all-opti}
sum_opti_all %>% 
  ggplot(aes(strategy, mem_gb_median, color = method)) +
  geom_point() +
  facet_wrap(dataset ~ ., scales = 'free_y') +
  scale_color_manual(values = tri_colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

```{r metrics_all-human, fig.width=10, fig.height=3.5}
all_human %>% 
  pivot_longer(c(mcc, sec, mem_gb, fraction_mapped), names_to = 'metric') %>% 
  ggplot(aes(strategy, value, color = tool)) +
  geom_boxplot() +
  facet_wrap('metric', scales = 'free_y', nrow=1) +
  labs(x='', y='',  
       caption = 'Only the human dataset is shown here. The patterns are similar with other datasets.') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
